CrypTool Script\footnote{ Background reading for the CrypTool program}\\ Mathematics and Cryptography

### 2003-06-11 script/de/chapters/digitalsignatures.tex
ecc + ds changes

Jeder Teilnehmer, der sich per \index{X.509} X.509-Zertifikat die Zugeh"origkeit seines "offentlichen Schl"ussels zu seiner
realen Person best"atigen lassen m"ochte, wendet sich an eine sogenannte \index{Certification Authority (CA)} Certification Authority (CA). Dieser
beweist er seine Identit"at (etwa durch Vorlage seines Personalausweises). Anschlie"send stellt die CA ihm ein
elektronisches Dokument (Zertifikat) aus, in dem im wesentlichen die Namen des Zertifikatnehmers und
der CA, der "offentliche Schl"ussel des Zertifikatnehmers und der G"ultigkeitszeitraum des Zertifikats vermerkt
sind. Die CA unterzeichnet das Zertifikat anschlie"send mit ihrem geheimen Schl"ussel.

@@

\bibitem[FIPS186-2]{FIPS186-2} U.S. Department of Commerce/N.I.S.T. 
    \index{FIPS186-2} \\
    Digital Signature Standard (DSS), January 27, 2000.\\
    Change Note: October 5, 2001.\\
    http://csrc.nist.gov/publications/fips/fips186-2/fips186-2-change1.pdf

>>>>

Jeder Teilnehmer, der sich per \index{X.509} X.509-Zertifikat \cite{X.509} die
Zugeh"origkeit seines "offentlichen Schl"ussels zu seiner realen Person
best"atigen lassen m"ochte, wendet sich an eine sogenannte \index{Certification
Authority (CA)} Certification Authority (CA). Dieser beweist er seine
Identit"at (etwa durch Vorlage seines Personalausweises). Anschlie"send stellt
die CA ihm ein elektronisches Dokument (Zertifikat) aus, in dem im wesentlichen
die Namen des Zertifikatnehmers und der CA, der "offentliche Schl"ussel des
Zertifikatnehmers und der G"ultigkeitszeitraum des Zertifikats vermerkt sind.
Die CA unterzeichnet das Zertifikat anschlie"send mit ihrem geheimen
Schl"ussel.

@@

\bibitem[FIPS186-2]{FIPS186-2} U.S. Department of Commerce/N.I.S.T. 
    \index{FIPS186-2} \\
    Digital Signature Standard (DSS), January 27, 2000.\\
    Change Note: October 5, 2001.\\
    http://csrc.nist.gov/publications/fips/fips186-2/fips186-2-change1.pdf
\bibitem[X.509]{X.509}
    \index{X.509}
    ITU-T Recommendation X.509 (1997 E): Information Technology -- 
    Open Systems Interconnection -- The Directory: Authentication Framework, June 1997.

### 2003-06-11 script/en/chapters/digitalsignatures.tex
ecc + ds changes

% --------------------------------------------------------------------------
\subsubsection{Impersonation attacks}
\index{Impersonation attack}
Assume Charlie has two pairs of keys (PK1, SK1) and (PK2, SK2), where SK denotes
the secret key and PK the public key. Further assume that he manages to palm off
PK1 on Alice as Bob's public key and PK2 on Bob as Alice's public key (by
falsifying a public key directory).

@@

% --------------------------------------------------------------------------
\subsubsection{X.509}
All participants who want to \index{X.509} have a X.509 certificate verifying that their public
key belongs to a real person consult what is known as a
\index{Certification authority (CA)} certification authority (CA). They prove their identity to this
CA (for example by showing their ID). The CA then issues them an electronic
document (certificate) which essentially contains the names of the certificate-holder
and the CA, the certificate-holder's public key and the validity period
of the certificate. The CA then signs the certificate using its secret key.

>>>>

% --------------------------------------------------------------------------
\subsubsection{Impersonation attacks}\label{Impersonalisierungsattacke}
\index{Impersonation attack}
Assume Charlie has two pairs of keys (PK1, SK1) and (PK2, SK2), where SK denotes
the secret key and PK the public key. Further assume that he manages to palm off
PK1 on Alice as Bob's public key and PK2 on Bob as Alice's public key (by
falsifying a public key directory).

@@

% --------------------------------------------------------------------------
\subsubsection{X.509}
All participants who want to \index{X.509} have a X.509 certificate
\cite{X.509} verifying that their public key belongs to a real person consult
what is known as a \index{Certification authority (CA)} certification authority
(CA). They prove their identity to this CA (for example by showing their ID).
The CA then issues them an electronic document (certificate) which essentially
contains the names of the certificate-holder and the CA, the
certificate-holder's public key and the validity period of the certificate. The
CA then signs the certificate using its secret key.

### 2003-06-11 script/en/chapters/ellipticcurves.tex
ecc + ds changes

Elliptic curves are used differently in the various areas. Encryption procedures
based on elliptic curves are based on the difficulty of a problem known as
elliptic curve discrete logarithm\index{Logarithm problem!discrete}. The factorisation of whole numbers uses the
fact that a large number of elliptic curves can be generated for a natural
composite number $n$ with several prime factors; however, these curves are not
then groups for composite $n$. \hyperlink{faktell}{More information about this
can be found under Factorisation using elliptic curves.}

@@

\subsection{Elliptic curves in cryptography}

In cryptography elliptic curve are a useful tool. Such curves are described by some equation. A detailed analysis has shown that curves of the form\footnote{This curve is given by the zeros of a {\it polynomial}\index{Polynomial} $F$ of degree three in three variables. In general, expressions of the form
$P=\sum_{i_1,\dots,i_n\in\N_0} a_{i_1\dots i_n} x_1^{i_1}\dots x_n^{i_n}$ with coefficients $a_{i_1\dots i_n}\in K$ are called polynomials in $n$ variables $x_1,\dots,x_n$ with underlying field $K$, if ${\rm deg\,} P:=\max\{i_1+\dots +i_n: a_{i_1\dots i_n}\ne 0\}$ is finite, i.e. the sum has only finitely many non-zero terms (monomials).}
\begin{equation}
 F(x_1,x_2,x_3)=-x_1^3+x_2^2x_3+a_1x_1x_2x_3-a_2x_1^2x_3+a_3x_2x_3^2-a_4x_1x_3^2-a_6x_3^3=0,
\label{eccbasisgleichung}
\end{equation}
are especially useful. The variables $x_1,x_2,x_3$ and parameters $a_1,\dots,a_4,a_6$ are elements of a given field $K$, which has certain properties that are make it useful from the cryptographic point of view. The underlying field $K$ might be the well known field of real numbers or some finite field (see last section).
In order to obtain a cure that is useful for cryptography, the parameters have to be chosen in a way that the following conditions hold
$$ {\partial F\over\partial x_1}\ne 0, \quad {\partial F\over\partial x_2}\ne 0, \quad
{\partial F\over\partial x_3}\ne 0 .
$$
We identify points on the curve that can be derived from each over by multiplying each component with some scalar. This makes sense since $(x_1,x_2,x_3)$ solves (\ref{eccbasisgleichung}) if and only if $\alpha (x_1,x_2,x_3)$ ($\alpha\ne 0$) does. Formally, this means that we consider classes of equivalent points instead of single points, where points are called equivalent if one is the scalar multiple of the other one.
\\ If we put $x_3=0$ in the basic equation (\ref{eccbasisgleichung}), then this equation collapses to $-x_1^3=0$, leading to $x_1=0$. Thus, the equivalence class which includes the element $(0,1,0)$ is the only one that contains a point with $x_3=0$. For all points on the curve that are not equivalent to $(0,1,0)$, we may apply the following transformation
$$ K\times K\times (K\setminus\{0\})\ni (x_1,x_2,x_3) \mapsto (x,y):=\left( {x_1\over x_3}, {x_2\over x_3}\right) \in K\times K \, ,
$$
which reduces the number of variables to two instead of three. We note that the basic equation (\ref{eccbasisgleichung}) $F(x_1,x_2,x_3)=0$  was chosen in a way that this transformation leads to the famous so-called Weierstrass-Equation\footnote{Karl Weierstrass, 1815-1897, German mathematician, famous for his rigorous formal approach to mathematics.} holds
\begin{equation}
 y^2+a_1xy+a_3y = x^3+a_2x^2+a_4x+a_6 \, .
\label{ell}
\end{equation}
Since all but one point (i.e. equivalence class) of the elliptic curve can be described using equation (\ref{ell}), this equation is often called the elliptic equation, and its solutions written as
$$ {\bf E} = \left\{(x,y)\in K\times K \, |\, y^2+a_1xy+a_3y = x^3+a_2x^2+a_4x+a_6  \right\} \cup \{{\cal O} \}.
$$
Here, ${\cal O}$ represents the point $(0,1,0)$ that is loosely speaking mapped to infinity by the transformation (division by $x_3$) that reduces the three variables to two.

@@

In order to prohibit a ``Man-in-the-middle"{}~attack, one may sign the values $G,q,R_A,R_B$ as described in chapter XXXX\ref{Impersonalisierungsattacke}.

@@

\hypertarget{faktell}{}

@@

Today it is still not clear whether these standardization initiatives will be successful or not. However, without useful standards, ECC is not likely to become a real alternative for RSA.

@@

Today elliptic curve cryptography is already in use. A prominent example is the information network Bonn-Berlin\footnote{Informationsverbund Bonn-Berlin (IVBB)}, used for the exchange of strictly confidential documents between different German federal governmental institutions in Berlin and Bonn. With the help of ECC a high security solutions could be realized. Interoperability, however, played only a minor role.

Based on information from the head of the Austrian e-Government projects, Prof.\ Posch, a smartcard based on ECC will shortly be launched in Austria: A bank card that allows digital signing will be issued in Austria from 2004 on.

@@

\bibitem[Lenstra1987]{Lenstra1987} H.W. Lenstra\\ \index{Lenstra 1987}
   Factoring integers with elliptic curves, 
   Annals of Mathematics 126, pp. 649-673, 1987.
\end{thebibliography}

>>>>

Elliptic curves are used differently in the various areas. Encryption
procedures based on elliptic curves are based on the difficulty of a problem
known as elliptic curve discrete logarithm\index{Logarithm problem!discrete}.
The factorisation of whole numbers uses the fact that a large number of
elliptic curves can be generated for a natural composite number $n$ with
several prime factors; however, these curves are not then groups for composite
$n$. More information about this can be found in chapter \ref{faktell}.

@@

\subsection{Elliptic curves in cryptography}\label{ECC-Crypto}

In cryptography elliptic curve are a useful tool. Such curves are described by some equation. A detailed analysis has shown that curves of the form\footnote{This curve is given by the zeros of a {\it polynomial}\index{Polynomial} $F$ of degree three in three variables. In general, expressions of the form
$P=\sum_{i_1,\dots,i_n\in\N_0} a_{i_1\dots i_n} x_1^{i_1}\dots x_n^{i_n}$ with coefficients $a_{i_1\dots i_n}\in K$ are called polynomials in $n$ variables $x_1,\dots,x_n$ with underlying field $K$, if ${\rm deg\,} P:=\max\{i_1+\dots +i_n: a_{i_1\dots i_n}\ne 0\}$ is finite, i.e. the sum has only finitely many non-zero terms (monomials).}
\begin{equation}
 F(x_1,x_2,x_3)=-x_1^3+x_2^2x_3+a_1x_1x_2x_3-a_2x_1^2x_3+a_3x_2x_3^2-a_4x_1x_3^2-a_6x_3^3=0,
\label{eccbasisgleichung}
\end{equation}
are especially useful. The variables $x_1,x_2,x_3$ and parameters $a_1,\dots,a_4,a_6$ are elements of a given field $K$, which has certain properties that are make it useful from the cryptographic point of view. The underlying field $K$ might be the well known field of real numbers or some finite field (see last section).
In order to obtain a cure that is useful for cryptography, the parameters have to be chosen in a way that the following conditions hold
$$ {\partial F\over\partial x_1}\ne 0, \quad {\partial F\over\partial x_2}\ne 0, \quad
{\partial F\over\partial x_3}\ne 0 .
$$
We identify points on the curve that can be derived from each over by multiplying each component with some scalar. This makes sense since $(x_1,x_2,x_3)$ solves (\ref{eccbasisgleichung}) if and only if $\alpha (x_1,x_2,x_3)$ ($\alpha\ne 0$) does. Formally, this means that we consider classes of equivalent points instead of single points, where points are called equivalent if one is the scalar multiple of the other one.
\\ If we put $x_3=0$ in the basic equation (\ref{eccbasisgleichung}), then this equation collapses to $-x_1^3=0$, leading to $x_1=0$. Thus, the equivalence class which includes the element $(0,1,0)$ is the only one that contains a point with $x_3=0$. For all points on the curve that are not equivalent to $(0,1,0)$, we may apply the following transformation
$$ K\times K\times (K\setminus\{0\})\ni (x_1,x_2,x_3) \mapsto (x,y):=\left( {x_1\over x_3}, {x_2\over x_3}\right) \in K\times K \, ,
$$
which reduces the number of variables to two instead of three. We note that the
basic equation (\ref{eccbasisgleichung}) $F(x_1,x_2,x_3)=0$  was chosen in a
way that this transformation leads to the famous so-called
Weierstrass-Equation\footnote{Karl Weierstrass, 31.10.1815$-$19.12.1897, German
mathematician, famous for his rigorous formal approach to mathematics.} holds
\begin{equation}
 y^2+a_1xy+a_3y = x^3+a_2x^2+a_4x+a_6 \, .
\label{ell}
\end{equation}
Since all but one point (i.e. equivalence class) of the elliptic curve can be described using equation (\ref{ell}), this equation is often called the elliptic equation, and its solutions written as
$$ {\bf E} = \left\{(x,y)\in K\times K \, |\, y^2+a_1xy+a_3y = x^3+a_2x^2+a_4x+a_6  \right\} \cup \{{\cal O} \}.
$$
Here, ${\cal O}$ represents the point $(0,1,0)$ that is loosely speaking mapped to infinity by the transformation (division by $x_3$) that reduces the three variables to two.

@@

In order to prohibit a ``Man-in-the-middle"{}~attack, one may sign the values $G,q,R_A,R_B$ as described in chapter \ref{Impersonalisierungsattacke}.

@@

\label{faktell}

@@

Today it is still not clear whether these standardization initiatives will be
successful or not. However, without agreed standards, ECC is not likely to
become a real alternative for RSA. The committees might be forced to act fast
if there was a break-through in factorization.

@@

Today elliptic curve cryptography is already in use. A prominent example is the information network Bonn-Berlin\footnote{Informationsverbund Bonn-Berlin (IVBB)}, used for the exchange of strictly confidential documents between different German federal governmental institutions in Berlin and Bonn. With the help of ECC a high security solution could be realized. Interoperability, however, played only a minor role.

Based on information from the head of the Austrian e-Government projects, Prof.\ Posch, a smartcard based on ECC will shortly be launched in Austria: A bank card that allows digital signing will be issued in Austria from 2004 on to all citizens.

@@

        \bibitem[Cassels1991]{Cassels1991} J. W. S. Cassels, {\em Lectures on elliptic curves}, Cambridge University Press, 1991, 143 Seiten.
	\bibitem[Koblitz1984]{Koblitz1984}N. Koblitz, {\em Introduction to elliptic curves and modular forms}, Graduate Texts in Mathemathics, Springer-Verlag, 1984.
	\bibitem[Menezes1993]{Menezes1993}A. J, Menezes, {\em Elliptic curve public key cryptosystems}, Kluwer Academic Publishers, 1993.
        \bibitem[Lenstra1987]{Lenstra1987} H.W. Lenstra \index{Lenstra 1987} \\
                Factoring integers with elliptic curves, Annals of Mathematics 126, pp. 649-673, 1987.
	\bibitem[Lenstra1999]{Lenstra1999} Arjen K. Lenstra, Eric R. Verheul  \index{Lenstra/Verheul 1999} \\
         	Selecting Cryptographic Key Sizes (1999), Journal of Cryptology: the journal of the International Association for Cryptologic Research \\
		\href{http://www.cryptosavvy.com/cryptosizes.pdf}{\texttt{http://www.cryptosavvy.com/cryptosizes.pdf}}
	\bibitem[Silverman1986]{Silverman1986} J. Silverman \\ The Arithmetic of Elliptic Curves, Springer-Verlag, 1986. \index{Silverman 1986}
	\bibitem[Silverman1992]{Silverman1992}J. Silverman, {\em The arithmetic of elliptc curves}, Graduate Texts in Mathemathics, Springer-Verlag, 1992.
	\bibitem[SivermanTate1992]{SivermanTate1992}J. Silverman, J. Tate, {\em Rational points on elliptic curves}, Springer-Verlag, 1992.

\end{thebibliography}
